Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-05-17 doc/GMP2015/revision_2/response_to_reviewers.txt

Response to Reviewers

=====================

Reviewer #1

-----------

Overall I think this paper is timely and interesting.  It is very nice that the authors have included a number of open questions to emphasize that there is more work to be done in this area.  The notation and terminology is rather dense, however, this is probably unavoidable given the intricacy of the questions being addressed.

@@

To that end, I would also suggest bringing the definition of geometric constraint system out of Appendix A and putting it right at the beginning of section 2, or perhaps section 2.1.

This line at the end of the abstract and at the end of section 2.1 should be modified for the final version of the paper (and applicable files should be included accordingly): ``We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.''

This paper seems essentially ready for publication.

RESPONSE TO REVIEWER # 1:

- Abstract had substantial rework. We felt it would be nontraditional to give a complete example in the abstract. But we do mention in the abstract which materials we model in the paper.

In the introduction, we explain much more clearly why we would want to use qusecs.

- We moved the definitions into the body of the paper.

- We cite a link where the video and code will be made available. We do not intend the code to be part of the submission. If the journal publishes videos linked to the paper, and the editors find it appropriate, we will upload the video files along with the final version.

Reviewer #2

-----------

This paper proposes a recursive decomposition planning algorithm to model a material layer as a solution of a 2D geometric constraint system. While the work seems important in material modeling, the topic is a kind of out of focus of GMP audiences. Besides, I'm not quite satisfied with the organization and structure of the paper, and I feel it is not suitable for publication in current form.

@@

1. The paper is far out of a 10 pages limit.
2. The basis concepts should be better put in the main context, but was put in an appendix (Appendix A).
3. The main algorithm in Section 4 should be stood out seperately and explained in details.
4. The examples should be enumerated.
5. The texts contain various formats and styles: Captioned, Bold fonts, colored, underlined, .... This unusual in scientific papers.

RESPONSE TO REVIEWER #2:

1. Our revision is being sent to the main journal rather than the
conference special issue and we were
told that the page limit does not apply.

2. We moved preliminaries into the main body of the paper (section 2).

3. We moved the algorithm into its own subsection and revised the proof to be more clear.

4. We have enumerated the examples.

5. We have addressed this and made the style uniform. We have reformatted our unusual headings etc..

Further Changes

===============

@@

Here we discuss additional changes the reviewers should be aware of.

After much discussion,  we are now submitting the revision to appear in the regular journal (see next section in this file) rather than the conference special issue.
The changes made in this revision are discussed below.

Title: Because of the substantial changes to the paper, we decided the title was no longer appropriately indicative of the revised paper.  Hence we have changed it to `Optimal Decomposition and Recombination of Isostatic Geometric Constraint Systems for Designing Layered Materials'. We believe this also makes the intent of our paper much more clear.

Throughout: Many minor corrections to spelling, grammar, and punctuation. Additionally, many sentences were rephrased/restructured to improve clarity. Also, because our paper is being considered to appear in the regular issue and does not have to adhere to the page limit, we have reformatted parts of the paper to use more space to improve readability and conform to standard formatting (addressing concerns of reviewer 2).

Some concepts were further fleshed out and discussed in this iteration:
(a) The comparison of our canonical DRP to the previously studied cluster-minimal DRP, and
(b) the reasons why neither canonical DRP nor cluster-minimal DRP's are optimal when the original constraint graph is overconstrained (not independent).
(c) We have also modified the writing so that it is clear that our theory extends to a larger class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition.

This is reflected throughout the paper, but is especially  in (1), (2), (3), and (8).

Abstract: Substantially altered for clarity.

@@

(1) Introduction: Entirely rewritten. A major complaint was that it was difficult to get a big picture understanding of the paper without reading the entire thing. We believe we addressed this with our revised intro (and abstract). This also includes a thorough explanation of `qusecs' and why we believe this notion is important to define and give a name, directly addressing reviewer criticism. Also, the discussion of previous work was moved into the corresponding sections (2) and (4).

(2) Preliminaries:  The definitions from the appendix (as per reviewer request) have been moved to this section.  Also reorganized the discussion of previous work on DR-plans.

(3) Canonical DR-plan: Improved the section introduction. Clarified/improved the definition of a canonical DR-plan. There are new figures to demonstrate the concepts; instead of the previous trivial (in the sense that it was tree-decomposable) example of three triangles, we use three $C2\times C3$s intersecting on a single triangle. Also, there is a new figure illustrating cluster minimality or canonical does not imply optimality in the case of overconstrained graphs. This was not present in the previous revision. We revised proofs for clarity, but there were no fundamental changes.

We put algorithm in a separate section (improved proof as mentioned) as per reviewer 2's request.  We moved the cluster minimality conjecture to the conclusion.  We added a discussion on the overconstrained problem and NP-hardness (3.3).

@@

(4) Recombination: Improved the section introduction. Moved in discussion of previous work on recombination from (1). Made the section that followed in the previous revision into a subsection (4.5).

(5) and (6) Application sections: No substantial changes other than rewordings.

(8) Conclusion: Moved in conjectures about MFA and cluster minimality from (3). Added new open problem (OP2).

@@

(App. A) Proofs: Revised DRP proofs for clarity, but there were no fundamental changes.

Journal Issue Change

==============

Originally the paper was submitted for consideration of publication in the special issue of CAGD for GMP2015.

However, for various reasons, this was changed to the regular issue of CAGD (and although our paper was invited to be presented at GMP15, we were unable to attend). We received confirmation from the GMP co-chairs that this paper would now be handled by the CAGD editors-in-chief.

Below, we have copied the email we received from the conference chairs, stating the change.

If there are additional questions in regard to this issue, please contact the corresponding author.

-----------

Reference email from GMP15 co-chairs

-----------

GMP2015 <gmp2015@easychair.org>

Wed, Mar 25, 2015 at 7:17 AM

To: Troy Baker <tbaker@cise.ufl.edu>

Dear Troy Baker,

### 2015-05-17 doc/GMP2015/revision_2/s01_introduction.tex
Revision 2, finished.

Geometric constraint systems have well-established, mature applications in mechanical engineering and  robotics, and they continue to find emerging applications in diverse fields from machine learning to molecular modeling.  Solving or realizing geometric constraint systems requires finding real solutions to a large multivariate polynomial system (of equalities and inequalities representing the constraints); this requires double exponential time in the number of variables, even if the type or orientation of the solution is specified.
% Thus, recursive decomposition into locally rigid subsystems (which have finitely many solutions) is crucial for realizing a geometric constraint system by the reverse process of recombining the subsystem solutions.
Thus, to realize a geometric constraint system, it is crucial to perform recursive decomposition into locally rigid subsystems (which have finitely many solutions), and then apply the reverse process of recombining the subsystem solutions.
With the use of \dfn{decomposition-recombination (DR-) planning}, the complexity is dominated by the size of the largest subsystem that is solved, or recombined, from the solutions of its child subsystems, i.e.\ the maximum fan-in occurring in a DR-plan.
In addition, navigating and analyzing the solution spaces, as well as designing constraint systems with desired solution spaces leads to the \dfn{optimal decomposition-recombination (DR-) planning} problem~\cite{sitharam2005combinatorial, hoffman2001decompositionI, hoffman2001decompositionII}. \todo{Feels very clunky.}

@@

\subsection{Introducing Qusecs}
A large class of constraint systems that we call \dfn{qusecs}, a contraction of ``quasi-uniform or self-similar constraint system'', (a) can be treated combinatorially as described above and (b) occur as independent (isostatic or underconstrained) systems in materials applications. We discuss these next.
%
Some natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers. As illustrated by the examples below, the structure within each layer is often: self-similar\footnote{In this manuscript we only study finite 2D structures. \dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.}~\cite{2012arXiv1204.6389G}, spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and composed of a few repeated motifs appearing in disordered arrangements.
Note that a 2D layer is not necessarily planar (genus 0), it can consist of multiple, inter-constraining planar monolayers. Furthermore, a layer is often  either \dfn{isostatic} or \dfn{underconstrained} (not \dfn{self-stressed}/\dfn{overconstrained}, \seedefsd). \todo{consider...} These properties, as well as quasi-uniformity, aperiodicity, self-similarity, and layered structure, are natural consequences of evolutionary pressures or design objectives such as stability, minimizing mass, optimally distributing external stresses, and participating in the assembly of diverse and multifunctional, larger structures.
%

@@

In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained qusecs are optimal. In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans when a system is over-constrained. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied {\em cluster minimality}  \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}). Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.  While this and the next section focus on bar-joint graphs, the theory is easily extended to other qusecs used to model the abovementioned types of materials, as shown in subsequent sections.
% that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
% Additionally, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.

@@

\todo{We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.}

>>>>

Geometric constraint systems have well-established, mature applications in mechanical engineering and  robotics, and they continue to find emerging applications in diverse fields from machine learning to molecular modeling.  Solving or realizing geometric constraint systems requires finding real solutions to a large multivariate polynomial system (of equalities and inequalities representing the constraints); this requires double exponential time in the number of variables, even if the type or orientation of the solution is specified.
% Thus, recursive decomposition into locally rigid subsystems (which have finitely many solutions) is crucial for realizing a geometric constraint system by the reverse process of recombining the subsystem solutions.
Thus, to realize a geometric constraint system, it is crucial to perform recursive decomposition into locally rigid subsystems (which have finitely many solutions), and then apply the reverse process of recombining the subsystem solutions.
With the use of \dfn{decomposition-recombination (DR-) planning}, the complexity is dominated by the size of the largest subsystem that is solved, or recombined, from the solutions of its child subsystems, i.e.\ the maximum fan-in occurring in a DR-plan.
In addition, navigating and analyzing the solution spaces, as well as designing constraint systems with desired solution spaces, leads to the \dfn{optimal decomposition-recombination (DR-) planning} problem~\cite{sitharam2005combinatorial, hoffman2001decompositionI, hoffman2001decompositionII}.

@@

\subsection{Introducing Qusecs}
A large class of constraint systems that we call \dfn{qusecs}, a contraction of ``quasi-uniform or self-similar constraint system'', (a) can be treated combinatorially as described above and (b) occur as independent (isostatic or underconstrained) systems in materials applications. We discuss these next.
%
Some natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers. As illustrated by the examples below, the structure within each layer is often: self-similar\footnote{In this manuscript we only study finite 2D structures. \dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.}~\cite{2012arXiv1204.6389G}, spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and composed of a few repeated motifs appearing in disordered arrangements.
Note that a 2D layer is not necessarily planar (genus 0), it can consist of multiple, inter-constraining planar monolayers. Furthermore, a layer is often  either \dfn{isostatic} or \dfn{underconstrained} (not \dfn{self-stressed}/\dfn{overconstrained}, \seedefsd). These properties, as well as quasi-uniformity, aperiodicity, self-similarity, and layered structure, are natural consequences of evolutionary pressures or design objectives such as stability, minimizing mass, optimally distributing external stresses, and participating in the assembly of diverse and multifunctional, larger structures.
%

@@

In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained qusecs are optimal. In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans when a system is over-constrained. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied \dfn{cluster minimality}~\cite{hoffman2001decompositionI}. Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.  While this and the next section focus on bar-joint graphs, the theory is easily extended to other qusecs used to model the abovementioned types of materials, as shown in subsequent sections.
% that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
% Additionally, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.

@@

A software implementation of our algorithms and videos demonstrating the software are publicly available online\footnote{See footnote 1.}.

### 2015-05-17 doc/GMP2015/revision_2/s02_preliminaries.tex
Revision 2, finished.

\subsection{Geometric Constraint systems and Combinatorial Rigidity}
\label{sec:prelim:defs}
\label{sec:appendix:defs}

@@

\subsection{Decomposition-Recombination (DR) Plans}
%
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{overconstrained} if it is not independent.
%
%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
%
%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
%
% \begin{definition}
%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
%
%
%
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%

@@

\ClearMyMinHeight
\SetMyMinHeight{.4}{img/svg/triangle}
\SetMyMinHeight{.3}{img/svg/triangle_candrp}
\SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}

@@

See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.

\subsection{Previous Work on DR-plans}
\label{sec:prev}
% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}
We now briefly survey existing techniques for detecting rigidity and creating DR-plans of 2D constraint systems. The limitations of these techniques directly motivate the contributions of the next  section.

\subsubsection{Finding (vertex)-maximal generically rigid subsystems}
Fast, graph-based algorithms exist (pebble-game~\cite{Jacobs:1997:PG,hoffmann1997solvablesubsets,jermann2006decomposition,Lee:2007:PGA}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs. When the input itself is rigid, these algorithms do nothing, i.e.\ compute the identity function.

@@

\subsubsection{Optimal Recursive Decomposition (DR-planning)}
Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{lomonosov2004graph,sitharam2005combinatorial,jermann2006decomposition} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim. For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm \todo{rephrase? maybe not} without involving geometric information in the constraint system. Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in~\cite{hoffman2001decompositionI}.
%An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system. Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.

@@

\subsubsection{DR-plans for special classes and with other criteria}
For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems~\cite{owen1991algebraic,fudos1997graph,joan-arinyo2004revisiting} common in computer aided mechanical design (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal. This satisfies the Church-Rosser property, leading to highly efficient DR-planning algorithms. For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to have a minimal set of at least 2 rigid proper subsystems as children (i.e.\ the union of no proper subset of size at least 2 child subsystems forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system. \seedefsc.

>>>>

\subsection{Geometric Constraint Systems and Combinatorial Rigidity}
\label{sec:prelim:defs}
\label{sec:appendix:defs}

@@

\subsection{Decomposition-Recombination (DR-) Plans}
%
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{overconstrained} if it is not independent.
%
%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
%
%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
%
% \begin{definition}
%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
%
%
%
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%

@@

% \ClearMyMinHeight
% \SetMyMinHeight{.4}{img/svg/triangle}
% \SetMyMinHeight{.3}{img/svg/triangle_candrp}
% \SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}

@@

See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.

\subsection{Previous Work on DR-Plans}
\label{sec:prev}
% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}
We now briefly survey existing techniques for detecting rigidity and creating DR-plans of 2D constraint systems. The limitations of these techniques directly motivate the contributions of the next  section.

\subsubsection{Finding (Vertex)-Maximal, Generically Rigid Subsystems}
Fast, graph-based algorithms exist (pebble-game~\cite{Jacobs:1997:PG,hoffmann1997solvablesubsets,jermann2006decomposition,Lee:2007:PGA}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs. When the input itself is rigid, these algorithms do nothing, i.e.\ compute the identity function.

@@

\subsubsection{Optimal Recursive Decomposition (DR-Planning)}
Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{lomonosov2004graph,sitharam2005combinatorial,jermann2006decomposition} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim. For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm without involving the geometric information in the constraint system. Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in~\cite{hoffman2001decompositionI}.
%An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system. Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.

@@

\subsubsection{DR-plans for Special Classes and with Other Criteria}
For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems~\cite{owen1991algebraic,fudos1997graph,joan-arinyo2004revisiting} common in computer aided mechanical design (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal. This satisfies the Church-Rosser property, leading to highly efficient DR-planning algorithms. For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to have a minimal set of at least 2 rigid proper subsystems as children (i.e.\ the union of no proper subset of size at least 2 child subsystems forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system. \seedefsc.
